Theory about Normal Basis on $F(2^k)$

### 2017-01-24 dissertation/submission/unsat.tex
dissertation: after submission_1st revision

\vspace{-0.5in}
\begin{center}
\begin{align*}
&\textit{Is\ subcircuit\ A\ functionally\ equivalent\ to\ subcircuit\ B?}\\
&\Longleftrightarrow
\textit{Is\ it\ true\ that\ no\ Boolean\ vector\ assignment\ to\ PIs\ a,b,c\ exists\ such\ that\ Z=1?}
\end{align*}
\end{center}

\begin{figure}[H]
\centering{
\includegraphics[scale=0.3]{newfig/fig_SAT.eps}}
\caption{An example of Boolean satisfiability problem on circuits.}
\label{fig:SAT}
\end{figure}

@@

On the other hand, researchers are seeking an alternative solution for SAT problem using a totally
different method from ``old-fashioned" DPLL algorithm. One promising option is polynomial calculus (PC),
mapping Boolean variables and connectors in CNF formulas to variables and operators in 
Galois fields. In this way clauses are transformed to monomials/polynomials, thus theorems and concepts in computer
algebra such as Hilbert's Nullstellensatz and Gr\"obner basis can be employed to assist finding
valid assignments or proofs of unsatisfiability. Basic concepts about PC will be formally introduced with definitions
from computer algebra.

\begin{table}[h]
\caption{Mapping Boolean operators to functions over $\Fkk$}
\centering
\begin{tabular}{|c|c|} 
\hline
 Boolean operator & Function over $\Fkk$ \\
\hline
\hline
 $a\land b$ & $a\cdot b$ \\
\hline
 $a\oplus b$ & $a+b$ \\
\hline
 $\bar{a}$ & $1+a$ \\
\hline
 $a\lor b$ & $a+b+a\cdot b$\\
\hline
\end{tabular}
\label{tab:booltof4}  
\end{table}

@@

\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
	\KwIn{$M$ -- original machine, $p$ -- property to check, $k$ -- \# of steps in $k$-BMC}
	\KwOut{If $p$ is violated, return error trace; otherwise $p$ is valid on $M$}
  $k = $ InitValue\;

  \eIf{$k$-BMC$(M,p,k)$ is \textbf{SAT}}
  {
	\Return{``Found error trace"}
  }
  {
	Extract UNSAT proof $\mathcal P$ of $k$-BMC\;
	$M' = $ \textit{ABSTRACT}$(M,\mathcal P)$\;
  }
  \eIf{MODEL-CHECK$(M',p)$ returns \textbf{PASS}}
  {
	\Return{``Passing property"}
  }
  {
	Increase bound $k$\;
	goto Line 2\;
  }
\caption {Abstraction refinement using $k$-BMC}\label{alg:absrefine}
\end{algorithm}
\DecMargin{1em}

\begin{figure}[h]
\centering{
\includegraphics[width=5.5in]{newfig/refine.eps}}
\caption{Abstraction by reducing latches.}
\label{fig:refine}
\end{figure}
Assume that we are given a sequential circuit with $n$ latches as shown in Figure \ref{fig:refine}(a). 
This circuit can be modeled as a Mealy machine $M$ and the states $s$ can be explicitly encoded by bit-level latch variables 
$l_1,\dots,l_n$. Algorithm \ref{alg:absrefine} describes an approach to check if machine $M$ violates property $p$.
This algorithm relies on $k$-BMC technique, which works on the basis of CNF-SAT solving.
The $k$-BMC represents the initial states $I$, the transition relation $T$ and property $p$ as CNF formulas.

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is SAT or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
SAT, then a counterexample is found that violates property $p$. If the result is UNSAT, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fix-point detection.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

@@

It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
the UNSAT core is extracted using a conventional CNF-SAT solver, which encounters the ``bit-blasting" problem
when the size of datapath (number of latches in Figure \ref{fig:refine}) is very large. \textbf{Here we propose an altogether new 
method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
method according to the following observation:}

@@

\begin{figure}[h]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}}
\caption{DAG representing $Spoly$ computations and multivariate divisions.}
\label{fig:originUNSAT}
\end{figure}

@@

\begin{algorithm}[hbt]
\SetAlgoNoLine
	\KwIn{A set of polynomials $F = \{f_1,f_2,\dots,f_s\}$}
	\KwOut{An UNSAT core $\{f_{m_1},f_{m_2},\dots,f_{m_t}\}$}
\Repeat{$r_l == 1$}
{
	Pick a pair $f_i,f_j\in F$ that has never been computed $Spoly$\;
	\If{$Spoly(f_i,f_j)\xrightarrow{F}_+ r_l \neq 0$}
	{
		$F = F\cup r_l$\;
		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the $Spoly$, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the $Spoly$\;
	}
}
Backward traverse the DAG for remainder ``1", replace $r_l$ with corresponding DAG $G_l$\;
\Return{All roots}
\caption {Extract UNSAT core using a variation of Buchberger's algorithm}\label{alg:UNSAT}
\end{algorithm}

@@

\begin{figure}[h]
\centering{
\includegraphics[width=4.5in]{newfig/refutation_tree.eps}}
\caption{Generating refutation trees to record UNSAT cores.}
\label{fig:refute}
\end{figure}

@@

After the first iteration of the GB-core algorithm, the core is
identified as $\{f_1, f_2,f_3,f_4\}$, and we obtain a
refutation tree as shown in Figure \ref{fig:refine}(a).  
\begin{figure}[h]
\centering{
\includegraphics[width=\textwidth]{newfig/core_refine.eps}}
\caption{Refutation trees of core refinement example.}
\label{fig:refine}
\end{figure}

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/s27.eps}}
\caption{Gate-level schematic of the example circuit.}
\label{fig:s27}
\end{figure}

Figure \ref{fig:s27} shows a sequential circuit (``s27'' from the
ISCAS benchmark set) with 3-bit state registers
PS = $\{G7,G6,G5\}$ and NS = $\{G13,G11,G10\}$. 
Its underlying FSM contains 8 states.

\begin{figure}[H]
\centering{
\includegraphics[width=4.5in]{newfig/s27_stg.pdf}}
\caption{State transition graph of example circuit.}
\label{fig:s27_stg}
\end{figure}

@@

The result shows that state variables $\{G5,G6, G10,G11\}$ are
irrelevant when considering the violation of $p$. Thus, we can remove
the latches $G10, G11$, and make $G5,G6$ primary inputs. In this way,
the number of state variables is reduced to 1, such that the new
machine only contains 2 states.

\begin{figure}[H]
\centerline{\includegraphics[width=4.5in]{newfig/s27_stg_refined.pdf}}
\caption{State transition graph of abstracted machine.}
\label{fig:s27_stg_refined}
\end{figure}

@@

%\vspace{-0.1in}
\begin{table}[htb]
\centering
\caption{Results of running benchmarks using our tool. 
\small{Asterisk($^*$) denotes that the benchmark was not translated from CNF.} 
Our tool is composed by 3 parts: part I runs a single GB-core algorithm,
part II applies the iterative refinement heuristic to run the GB-core algorithm
iteratively, part III applies the syzygy heuristic.}
{\small 
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
\hline
\multirow{3}{1.8cm}{\centering Benchmark} 
& \multirow{3}{0.7cm}{\centering \# Polys} 
& \multirow{3}{0.7cm}{\centering \# MUS} 
& \multicolumn{3}{c|}{\multirow{2}{1.5cm}{\centering Size of core}}
 & \multirow{3}{1.3cm}{\centering \# GB-core iterations}
 & \multicolumn{3}{c|}{\multirow{2}{1.8cm}{\centering Runtime (sec)}}
 & \multirow{3}{1.6cm}{\centering Runtime of PicoMUS (sec)} \\
  & & &\multicolumn{3}{c|}{}& &\multicolumn{3}{c|}{}& \\
  \cline{4-6} \cline{8-10}
    & & & I & II & III & & I & II & III & \\
\hline
\hline
$5\times 5$ SMPO & 240  & 137  & 169 & 137 & 137  & 8  & 1222 & 1938 & 1698 & $<0.1$\\
$4\times 4$ SMPO$^*$ & 84  & 21  & 21 & 21 & 21 & 1  & 125 & 0.3 & 29  & - \\
$3\times 3$ SMPO$^*$ & 45  & 15  & 15 & 15 & 15  & 1  & 6.6 & 0.2 & 5.7 & - \\
$3 \times 3$ SMPO & 17 & 2 & 2 &2 & 2 & 1 & 0.07 & 0.01 & 0.01  & $<0.1$  \\
$4 \times 4$ MasVMont$^*$ & 148 & 83 & 83 & 83 & 83 & 1 & 23 & 139 & 12 & - \\
$3 \times 3$ MasVMont$^*$ & 84 & 53 & 53 & 53 & 53  & 1 & 4.3 & 4.6 & 0.9  & - \\
$2 \times 2$ MasVMont & 27 & 23 & 24 & 23 & 23 & 2 & 1.3 & 1.0 & 80  & $<0.1$ \\
$5\times 5$ RH$^*$ & 142  & 34  & 48 & 35 & 35  & 4  & 997 & 1.0 & 80 & - \\
$4\times 4$ RH$^*$ & 104  & 35  & 43 & 36 & 36  & 3  & 96 & 5.7 & 0.6 & -\\
$3\times 3$ RH$^*$ & 50  & 20  & 20 & 20 & 20  & 1  &2.9 & 3.5 & 10  & -\\
aim-100 & 79 & 22 & 22 & 22 & 22 & 1  & 43 & 0.7 & 0.2 & $<0.1$\\
cocktail & 135 & 4 & 6 & 4 & 4 & 2 & 51 & 0.01 & 0.01  & $<0.1$ \\
subset-1 & 100 & 6 & 6 & 6 & 6 & 1 & 2.4 & 0.01 & 0.01 & $<0.1$ \\
subset-2 & 141 & 19 & 37 & 23 & 21 & 2 & 12 & 1.6 & 1.1 & $<0.1$  \\
subset-3 & 118 & 16 & 13 & 12 & 11 & 2 & 8.6 & 0.2 & 0.07 & $<0.1$ \\
phole4 & 104 & 10 & 16 & 16 & 10 & 1 & 4.3 & 0.2 & 0.5 &  $<0.1$\\
phole5 & 169 & 19 & 30 & 25 & 19 & 3 & 12 & 3.2 & 2.7 & $<0.1$ \\
\hline
\end{tabular}
}
\label{tab:result}  
\end{table}

>>>>

\vspace{-0.7in}
\begin{center}
\begin{align*}
&\textit{Is\ subcircuit\ A\ functionally\ equivalent\ to\ subcircuit\ B?}\\
&\Longleftrightarrow
\textit{Is\ it\ true\ that\ no\ Boolean\ vector\ assignment\ to\ PIs\ a,b,c\ exists\ such\ that\ Z=1?}
\end{align*}
\end{center}

\begin{figure}[bp]
\centering{
\includegraphics[scale=0.3]{newfig/fig_SAT.eps}}
\caption{An example of Boolean satisfiability problem on circuits.}
\label{fig:SAT}
\end{figure}

@@

On the other hand, researchers are seeking an alternative solution for SAT problem using a totally
different method from ``old-fashioned" DPLL algorithm. One promising option is polynomial calculus (PC),
mapping Boolean variables and connectors in CNF formulas to variables and operators in 
Galois fields as examples in Table \ref{tab:booltof4}. In this way clauses are transformed to monomials/polynomials, thus theorems and concepts in computer
algebra such as Hilbert's Nullstellensatz and Gr\"obner basis can be employed to assist finding
valid assignments or proofs of unsatisfiability. Basic concepts about PC will be formally introduced with definitions
from computer algebra.

\begin{table}[bp]
\caption{Mapping Boolean operators to functions over $\Fkk$}
\centering
\begin{tabular}{|c|c|} 
\hline
 Boolean operator & Function over $\Fkk$ \\
\hline
\hline
 $a\land b$ & $a\cdot b$ \\
\hline
 $a\oplus b$ & $a+b$ \\
\hline
 $\bar{a}$ & $1+a$ \\
\hline
 $a\lor b$ & $a+b+a\cdot b$\\
\hline
\end{tabular}
\label{tab:booltof4}  
\end{table}

@@

% \IncMargin{1em}
\begin{algorithm}[htb]
\SetAlgoNoLine
\LinesNumbered
	\KwIn{$M$ -- original machine, $p$ -- property to check, $k$ -- \# of steps in $k$-BMC}
	\KwOut{If $p$ is violated, return error trace; otherwise $p$ is valid on $M$}
  $k = $ InitValue\;

  \eIf{$k$-BMC$(M,p,k)$ is \textbf{SAT}}
  {
	\Return{``Found error trace"}
  }
  {
	Extract UNSAT proof $\mathcal P$ of $k$-BMC\;
	$M' = $ \textit{ABSTRACT}$(M,\mathcal P)$\;
  }
  \eIf{MODEL-CHECK$(M',p)$ returns \textbf{PASS}}
  {
	\Return{``Passing property"}
  }
  {
	Increase bound $k$\;
	goto Line 2\;
  }
\caption {Abstraction refinement using $k$-BMC}\label{alg:absrefine}
\end{algorithm}
% \DecMargin{1em}

Assume that we are given a sequential circuit with $n$ latches as shown in Figure \ref{fig:absrefine}(a). 
This circuit can be modeled as a Mealy machine $M$ and the states $s$ can be explicitly encoded by bit-level latch variables 
$l_1,\dots,l_n$. Algorithm \ref{alg:absrefine} describes an approach to check if machine $M$ violates property $p$.
This algorithm relies on $k$-BMC technique, which works on the basis of CNF-SAT solving.
The $k$-BMC represents the initial states $I$, the transition relation $T$ and property $p$ as CNF formulas.

\begin{figure}[bp]
\centering{
\includegraphics[width=5.5in]{newfig/refine.eps}}
\caption{Abstraction by reducing latches.}
\label{fig:absrefine}
\end{figure}

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is SAT or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
SAT, then a counterexample is found that violates property $p$. If the result is UNSAT, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fix-point detection.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:absrefine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

@@

It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
the UNSAT core is extracted using a conventional CNF-SAT solver, which encounters the ``bit-blasting" problem
when the size of datapath (number of latches in Figure \ref{fig:absrefine}) is very large. \textbf{Here we propose an altogether new 
method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
method according to the following observation:}

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}}
\caption{DAG representing $Spoly$ computations and multivariate divisions.}
\label{fig:originUNSAT}
\end{figure}

@@

\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
	\KwIn{A set of polynomials $F = \{f_1,f_2,\dots,f_s\}$}
	\KwOut{An UNSAT core $\{f_{m_1},f_{m_2},\dots,f_{m_t}\}$}
\Repeat{$r_l == 1$}
{
	Pick a pair $f_i,f_j\in F$ that has never been computed $Spoly$\;
	\If{$Spoly(f_i,f_j)\xrightarrow{F}_+ r_l \neq 0$}
	{
		$F = F\cup r_l$\;
		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the $Spoly$, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the $Spoly$\;
	}
}
Backward traverse the DAG for remainder ``1", replace $r_l$ with corresponding DAG $G_l$\;
\Return{All roots}
\caption {Extract UNSAT core using a variation of Buchberger's algorithm}\label{alg:UNSAT}
\end{algorithm}

@@

\begin{figure}[tbp]
\centering{
\includegraphics[width=4.5in]{newfig/refutation_tree.eps}}
\caption{Generating refutation trees to record UNSAT cores.}
\label{fig:refute}
\end{figure}

@@

After the first iteration of the GB-core algorithm, the core is
identified as $\{f_1, f_2,f_3,f_4\}$, and we obtain a
refutation tree as shown in Figure \ref{fig:refine}(a).  
\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/core_refine.eps}}
\caption{Refutation trees of core refinement example.}
\label{fig:refine}
\end{figure}

@@

\begin{Example}
Figure \ref{fig:s27} shows a sequential circuit (``s27'' from the
ISCAS benchmark set) with 3-bit state registers
PS = $\{G7,G6,G5\}$ and NS = $\{G13,G11,G10\}$. 
Its underlying FSM contains 8 states, as shown in Figure \ref{fig:s27_stg}

\begin{figure}[tbp]
\centering{
\includegraphics[width=\textwidth]{newfig/s27.eps}}
\caption{Gate-level schematic of the example circuit.}
\label{fig:s27}
\end{figure}

\begin{figure}[tbp]
\centering{
\includegraphics[width=4.5in]{newfig/s27_stg.pdf}}
\caption{State transition graph of example circuit.}
\label{fig:s27_stg}
\end{figure}

@@

The result shows that state variables $\{G5,G6, G10,G11\}$ are
irrelevant when considering the violation of $p$. Thus, we can remove
the latches $G10, G11$, and make $G5,G6$ primary inputs. In this way,
the number of state variables is reduced to 1, such that the new
machine only contains 2 states, as the STG in Figure \ref{fig:s27_stg_refined}.

\begin{figure}[tbp]
\centerline{\includegraphics[width=4.5in]{newfig/s27_stg_refined.pdf}}
\caption{State transition graph of abstracted machine.}
\label{fig:s27_stg_refined}
\end{figure}

@@

%\vspace{-0.1in}
\begin{table}[tbp]
\centering
\caption{Results of running benchmarks using our tool. 
\small{Asterisk($^*$) denotes that the benchmark was not translated from CNF.} 
Our tool is composed by 3 parts: part I runs a single GB-core algorithm,
part II applies the iterative refinement heuristic to run the GB-core algorithm
iteratively, part III applies the syzygy heuristic.}
{\small 
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
\hline
\multirow{3}{1.8cm}{\centering Benchmark} 
& \multirow{3}{0.7cm}{\centering \# Polys} 
& \multirow{3}{0.7cm}{\centering \# MUS} 
& \multicolumn{3}{c|}{\multirow{2}{1.5cm}{\centering Size of core}}
 & \multirow{3}{1.3cm}{\centering \# GB-core iterations}
 & \multicolumn{3}{c|}{\multirow{2}{1.8cm}{\centering Runtime (sec)}}
 & \multirow{3}{1.6cm}{\centering Runtime of PicoMUS (sec)} \\
  & & &\multicolumn{3}{c|}{}& &\multicolumn{3}{c|}{}& \\
  \cline{4-6} \cline{8-10}
    & & & I & II & III & & I & II & III & \\
\hline
\hline
$5\times 5$ SMPO & 240  & 137  & 169 & 137 & 137  & 8  & 1222 & 1938 & 1698 & $<0.1$\\
$4\times 4$ SMPO$^*$ & 84  & 21  & 21 & 21 & 21 & 1  & 125 & 0.3 & 29  & - \\
$3\times 3$ SMPO$^*$ & 45  & 15  & 15 & 15 & 15  & 1  & 6.6 & 0.2 & 5.7 & - \\
$3 \times 3$ SMPO & 17 & 2 & 2 &2 & 2 & 1 & 0.07 & 0.01 & 0.01  & $<0.1$  \\
$4 \times 4$ MasVMont$^*$ & 148 & 83 & 83 & 83 & 83 & 1 & 23 & 139 & 12 & - \\
$3 \times 3$ MasVMont$^*$ & 84 & 53 & 53 & 53 & 53  & 1 & 4.3 & 4.6 & 0.9  & - \\
$2 \times 2$ MasVMont & 27 & 23 & 24 & 23 & 23 & 2 & 1.3 & 1.0 & 80  & $<0.1$ \\
$5\times 5$ RH$^*$ & 142  & 34  & 48 & 35 & 35  & 4  & 997 & 1.0 & 80 & - \\
$4\times 4$ RH$^*$ & 104  & 35  & 43 & 36 & 36  & 3  & 96 & 5.7 & 0.6 & -\\
$3\times 3$ RH$^*$ & 50  & 20  & 20 & 20 & 20  & 1  &2.9 & 3.5 & 10  & -\\
aim-100 & 79 & 22 & 22 & 22 & 22 & 1  & 43 & 0.7 & 0.2 & $<0.1$\\
cocktail & 135 & 4 & 6 & 4 & 4 & 2 & 51 & 0.01 & 0.01  & $<0.1$ \\
subset-1 & 100 & 6 & 6 & 6 & 6 & 1 & 2.4 & 0.01 & 0.01 & $<0.1$ \\
subset-2 & 141 & 19 & 37 & 23 & 21 & 2 & 12 & 1.6 & 1.1 & $<0.1$  \\
subset-3 & 118 & 16 & 13 & 12 & 11 & 2 & 8.6 & 0.2 & 0.07 & $<0.1$ \\
phole4 & 104 & 10 & 16 & 16 & 10 & 1 & 4.3 & 0.2 & 0.5 &  $<0.1$\\
phole5 & 169 & 19 & 30 & 25 & 19 & 3 & 12 & 3.2 & 2.7 & $<0.1$ \\
\hline
\end{tabular}
}
\label{tab:result}  
\end{table}
